---
title: "Prepare Dataset for CDI Risk Assessment Model"
author: "Steph Reynolds"
date: "`r format(Sys.time(), '%b %d, %Y  %H:%M %p')`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Background

## Project

MINDSCAPE: Modeling of infectious network dynamics for surveillance, control and prevention enhancement

## Description

This file reads in necessary data and identifies variables of interest for building a multivariate logistic regression model to evaluate C diff risk among hospitalized patients. The resulting data frame (`df.CDI`) contains binary indicators for CDI risk factors. Variables include:

-   `ID` --\> each patient's unique de-identified encounter ID
-   `readmit` --\> whether patient had a prior hospital admission in the last 30 days
-   `ED_dispo` --\> whether pt was admitted from Emergency Dept (ED) ???
    -   NOTE: UNCLEAR! VERIFY / LEO WHAT THIS SHOULD BE TO INDICATE THAT PT WAS ADMITTED TO ED!
-   `age_gte_65` -\> whether pt is aged 65 years or older at time of hospitalization
-   `PPI` --\> whether pt received proton pump inhibitor (PPI) during hospital stay
-   `GAS` --\> whether pt received gastric acid suppressants (GAS) during hospital stay
    -   NOTE: This includes "ANTACIDS" and "HISTAMINE H2-RECEPTOR INHIBITORS"
-   `anyGAS` --\> whether pt received PPI and/or GAS during hospital stay
-   `LAX` --\> whether pt received laxatives (LAX) during hospital stay
    -   NOTE: This includes "LAXATIVES AND CATHARTICS" and "LAXATIVES, LOCAL/RECTAL"
-   `ABX` --\> whether pt received antibiotics during hospital stay
    -   For pts with a positive CDIF culture, only count ABX received up to day prior to `order_time`. For pts without CDI, count abx from time of admission to discharge.
-   `AMR` -\> whether pt received antimicrobials during hospital stay
    -   For pts with a positive CDIF culture, only count AMR received up to day prior to `order_time`. For pts without CDI, count abx from time of admission to discharge.
-   `DBM` --\> whether pt is diagnosed with diabetes mellitus (DBM)
    -   NOTE: DO NOT USE! ONLY HAVE DBM DX FOR COVID-POSITIVE PTS.

# Load required packages

```{r Load required packages}
library(tidyverse)
library(lubridate)
library(tableone) # create categorical tables
library(DescTools) # %like% operator
```

# Import `dm`, `med`, and `lab` tables and select only necessary variables

```{r Import data, message = F}
dm <- read_delim("/Users/sreynolds2/Downloads/ucsf_data_pull_11.08.21/demographics and event table 11.08.21.csv", delim = "|")

#med <- read_delim("/Users/sreynolds2/Downloads/ucsf_data_pull_11.08.21/med admin table 11.08.21.csv", delim = "|") %>% select(deid_enc_id, name, thera_class, pharm_class, taken_time)

lab <- read_delim("/Users/sreynolds2/Downloads/ucsf_data_pull_11.08.21/lab results 11.08.21.csv", delim = "|") %>% select(deid_enc_id, COMPONENT_ID, Abnormal, ORD_VALUE, order_time, RESULT_TIME, lab_concept)

dx <- read_delim("/Users/sreynolds2/Downloads/ucsf_data_pull_11.08.21/covid diagnoses 11.08.21.csv", delim = "|") %>% select(deid_enc_id, DX_NAME, dx_group)
```

# Clean `lab` table and identify all patients who had an abnormal CDIF test

```{r Clean `lab` table}
# Create not-like operator
`%notlike%` <- Negate(`%like%`)

# Create vector to exclude those where ORD_VALUE=='Test not performed'|'Indeterminate'
exclude_ord_value <- "Test not performed|Indeterminate"

# Identify all patients who had an abnormal CDIF test and assign to table `cdif_pts`
cdif_pts <- lab %>% 
  filter(COMPONENT_ID=='1301' & 
           Abnormal=='Abnormal' & 
           ORD_VALUE %notlike% exclude_ord_value) %>% 
  group_by(deid_enc_id) %>% 
  filter(order_time==min(order_time)) %>% 
  ungroup() %>% 
  distinct(deid_enc_id, order_time, RESULT_TIME) %>% 
  rename(CDIF_order_time = order_time,
         CDIF_result_time = RESULT_TIME)
```

# Join `dm` and `cdif_pts` tables and create new vars to indicate age\>=65, readmit status, and hospital-acquired CDIF

**\*Note:** Hospital-acquired CDIF (`HACDIF`) is defined as those with CDIF orders (`CDIF_order_time`) more than 48 hours after admission (`admit_time`).

```{r Join `dm` and `cdif_pts` tables}
df <- dm %>% 
  full_join(cdif_pts, by = "deid_enc_id") %>% # Join `dm` with `cdif_pts` table
  select(deid_enc_id, age, readmit, ED_dispo, admit_time, CDIF_order_time, CDIF_result_time, discharge_time) %>%
  mutate_at(c("admit_time", "CDIF_order_time", "discharge_time"), ymd_hms) %>% 
  mutate_at(c("readmit", "ED_dispo"), factor) %>%
  mutate(age_gte_65 = if_else(age>=65, 1, 0),
         readmit = if_else(readmit=="Yes", 1, 0),
         HACDIF = case_when((difftime(CDIF_order_time, admit_time, units = "hours")) > 48 ~ 1, TRUE ~ 0))

# Calculate proportion of encounter IDs with HACDIF 
cat("Proportion of total encounter IDs with hospital-acquired CDIF:",
    round(sum(df$HACDIF)/length(df$deid_enc_id), 4))
```

# Create binary variables to indicate whether patients received PPI, GAS, LAX, and/or ABX; assign to table `med2`

```{r Create binary vars to indicate whether pts received PPI, GAS, LAX, ABX, eval = F}
# Vector of antibiotic names - not captured by filtering thera_class=='ANTIBIOTICS' since thera_class=='NULL'
names_abx <- "VANCOMYCIN|CEFTRIAXONE|CEFEPIME|LEVOFLOXACIN|CIPROFLOXACIN|AMIKACIN|GENTAMICIN|NEOMYCIN-POLYMYXIN|TOBRAMYCIN|ERTAPENEM|IMIPENEM-CILASTATIN|MEROPENEM|MEROPENEM-VABORBACTAM|CEFAZOLIN|CEFOXITIN|CEFOTETAN|CEFUROXIME|CEFTAZIDIME|CEFTAROLINE|TOBRAMYCIN|DALBAVANCIN|ORITAVANCIN|CLINDAMYCIN|DAPTOMYCIN|AZITHROMYCIN|ERYTHROMYCIN|AZTREONAM|LINEZOLID|AMPICILLIN|AMPICILLIN-SULBACTAM|NAFCILLIN|PENICILLIN|PIPERACILLIN-TAZOBACTAM|BACITRACIN|NEOMYCIN-POLYMYXIN|GENTAMICIN-POLYMYXIN|TRIMETHOPRIM-SULFAMETHOXAZOLE|SULFAMETHOXAZOLE-TRIMETHOPRIM|DOXYCYCLINE|MINOCYCLINE|CLOFAZIMINE|ISONIAZID|RIFAMPIN|TIGECYCLINE|TRIMETHOPRIM|AZTREONAM|GENTAMICIN|CEFTOLOZANE-TAZOBACTAM|CLINDAMYCIN|ALBAVANCIN|ERAVACYCLINE|GENTAMICIN-POLYMYXIN|CEFAZOLIN/CEFTAZIDIME|ORITAVANCIN|PENICILLIN|TEDIZOID|CEFIDEROCOL|COLISTIMETHATE"

# Vector of antimicrobial names 
names_amr <- "VANCOMYCIN|CEFTRIAXONE|CEFEPIME|LEVOFLOXACIN|CIPROFLOXACIN|AMIKACIN|GENTAMICIN|NEOMYCIN|POLYMYXIN|TOBRAMYCIN|ERTAPENEM|IMIPENEM|CILASTATIN|MEROPENEM|MEROPENEM|VABORBACTAM|CEFAZOLIN|CEFOXITIN|CEFOTETAN|CEFUROXIME|CEFTAZIDIME|CEFTAROLINE|TOBRAMYCIN|DALBAVANCIN|ORITAVANCIN|CLINDAMYCIN|DAPTOMYCIN|AZITHROMYCIN|ERYTHROMYCIN|AZTREONAM|LINEZOLID|AMPICILLIN|AMPICILLIN|SULBACTAM|NAFCILLIN|PENICILLIN|PIPERACILLIN|TAZOBACTAM|BACITRACIN|NEOMYCIN|POLYMYXIN|GENTAMICIN|POLYMYXIN|TRIMETHOPRIM|SULFAMETHOXAZOLE|SULFAMETHOXAZOLE|TRIMETHOPRIM|DOXYCYCLINE|MINOCYCLINE|CLOFAZIMINE|ISONIAZID|RIFAMPIN|TIGECYCLINE|TRIMETHOPRIM|AZTREONAM|GENTAMICIN|CEFTOLOZANE|TAZOBACTAM|CLINDAMYCIN|ALBAVANCIN|ERAVACYCLINE|GENTAMICIN|POLYMYXIN|CEFAZOLIN/CEFTAZIDIME|ORITAVANCIN|PENICILLIN|TEDIZOID|CEFIDEROCOL|COLISTIMETHATE|ACYCLOVIR|LETERMOVIR|GANCICLOVIR|ACYCLOVIR|CIDOFOVIR|FOSCARNET|REMDESIVIR|PERAMIVIR|CASPOFUNGIN|POSACONAZOLE|AMPHOTERICIN|ISAVUCONAZONIUM|MICAFUNGIN|VORICONAZOLE"

# Create binary variables to indicate whether pt received PPI, GAS, anyGAS, LAX, ABX, and AMR
  # NOTE: This code takes ~ 10 minutes to run; consider putting into separate script
meds_binary <- med %>% 
  mutate_at(c("thera_class", "pharm_class"), factor) %>%
  mutate(PPI = if_else(pharm_class=="PROTON-PUMP INHIBITORS" | name %like% "PANTOPRAZOLE", 1, 0),    
         GAS = if_else(pharm_class %in% c("ANTACIDS", "HISTAMINE H2-RECEPTOR INHIBITORS"), 1, 0),
         anyGAS = if_else(pharm_class %in% c("PROTON-PUMP INHIBITORS", "ANTACIDS", "HISTAMINE H2-RECEPTOR INHIBITORS") |
                          name %like% "PANTOPRAZOLE", 1, 0),
         LAX = if_else(pharm_class %in% c("LAXATIVES AND CATHARTICS", "LAXATIVES, LOCAL/RECTAL"), 1, 0),
         ABX = if_else(thera_class=="ANTIBIOTICS" | name %like% names_abx, 1, 0),
         AMR = if_else(thera_class %in% c("ANTIBIOTICS","ANTIFUNGALS","ANTIINFECTIVES/MISCELLANEOUS",
                                          "ANTIPARASITICS","ANTIVIRALS") | name %like% names_amr, 1, 0)) %>% 
  select(-c(thera_class, pharm_class))

# Save as CSV and RDS file - that way, can just read in the file and don't have to keep re-running code 
write_csv(meds_binary, "/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/meds_binary.csv")
saveRDS(meds_binary, "/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/meds_binary.rds")
```

# Create `med_taken_before_HAI` table to indicate whether pt received PPI, GAS, ABX, or AMR up to one day before CDIF_order_time

**NOTE TO SELF:** Need to edit this section so that all ABX/PPI/GAS are counted, regardless of whether they were received up to one day before `CDIF_order_time`. Need to edit filter (line 3).

```{r Create `med_taken_before_HAI` table, eval = F}
# Read in `meds_binary` table 
meds_binary <- readRDS("/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/meds_binary.rds")

# For those with HA-CDIF, check whether they received any meds the calendar day before `CDIF_order_time`
meds_before_order <- df %>% 
  filter(HACDIF==1) %>%
  left_join(meds_binary, by = "deid_enc_id") %>% 
  mutate_at(c("CDIF_order_time", "taken_time"), as.Date) %>% 
  group_by(deid_enc_id) %>%
  mutate(PPI = if_else(sum(PPI)>=1, 1, 0),
         GAS = if_else(sum(GAS)>=1, 1, 0),
         anyGAS = if_else(sum(anyGAS)>=1, 1, 0),
         LAX = if_else(sum(LAX)>=1, 1, 0),
         ABX = if_else(sum(ABX)>=1, 1, 0),
         AMR = if_else(sum(AMR)>=1, 1, 0),
         taken_before_order = if_else(taken_time < CDIF_order_time, 1, 0))

# For those without CDIF culture, check whether they received any meds the calendar day before `CDIF `CDIF_order_time`
meds_before_disc <- df %>% 
  filter(is.na(CDIF_order_time)) %>%  # filter for missing order_time -- meaning pts with no CDIF order
  left_join(meds_binary, by = "deid_enc_id") %>% 
  group_by(deid_enc_id) %>% 
  mutate(PPI = if_else(sum(PPI)>=1, 1, 0),
         GAS = if_else(sum(GAS)>=1, 1, 0),
         anyGAS = if_else(sum(anyGAS)>=1, 1, 0),
         LAX = if_else(sum(LAX)>=1, 1, 0),
         ABX = if_else(sum(ABX)>=1, 1, 0),
         AMR = if_else(sum(AMR)>=1, 1, 0))

# Merge/bind rows, select binary med variables, and keep only unique encounters 
meds_before <- meds_before_disc %>% 
  rbind(meds_before_order) %>% 
  mutate(PPI = if_else(PPI==1 & (taken_before_order==1 | is.na(taken_before_order)), 1, 0),
         GAS = if_else(GAS==1 & (taken_before_order==1 | is.na(taken_before_order)), 1, 0),
         anyGAS = if_else(anyGAS==1 & (taken_before_order==1 | is.na(taken_before_order)), 1, 0),
         LAX = if_else(LAX==1 & (taken_before_order==1 | is.na(taken_before_order)), 1, 0),
         ABX = if_else(ABX==1 & (taken_before_order==1 | is.na(taken_before_order)), 1, 0),
         AMR = if_else(AMR==1 & (taken_before_order==1 | is.na(taken_before_order)), 1, 0)) %>% 
  select(deid_enc_id, PPI:AMR) %>% 
  distinct(deid_enc_id, .keep_all = T)

# Save as CSV and RDS file - that way, can just read in the file and don't have to keep re-running code 
write_csv(meds_before, "/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/meds_before_order_or_disc.csv")
saveRDS(meds_before, "/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/meds_before_order_or_disc.rds")
```

# Clean `dx` table and create binary variable to indicate whether patient has diabetes; assign to table `dbm`

```{r Clean `dx` table, create table to indicate DBM dx, eval = F}
dbm <- dx %>% 
  mutate(DBM = if_else(dx_group=="DIABETES MELLITUS", 1, 0)) %>%
  group_by(deid_enc_id) %>% 
  summarize(n_dbm = sum(DBM)) %>% 
  mutate(DBM = if_else(n_dbm>=1, 1, 0)) %>% 
  select(-n_dbm)
```

**NOTE:** Only have diabetes diagnoses for 1,210 out of 54,955 patients. Missing data on DBM diagnosis for \~97% of patients. Only have DBM data for Covid-positive patients. Checked demographics and events table, and diabetes dx not listed as variable. Would need Leo to re-pull data gathering diabetes dx for ALL patients.

# Join `df`, `med_taken_before_HACDIF`, and `dbm` tables and assign to `df_CDI`

```{r Merge all tables}
# Read in `meds_before_order_or_disc` 
meds_before <- readRDS("/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/meds_before_order_or_disc.rds")

# Updated Code from 3/31/2022
df_CDI <- df %>% 
  left_join(meds_before, by = 'deid_enc_id') %>%   # Left join dm and meds_before
  mutate(PPI = ifelse(is.na(PPI), 0, PPI),         # Reassign NA values for PPI:AMR to 0
         GAS = ifelse(is.na(GAS), 0, GAS),
         anyGAS = ifelse(is.na(anyGAS), 0, anyGAS),
         LAX = ifelse(is.na(LAX), 0, LAX),
         ABX = ifelse(is.na(ABX), 0, ABX), 
         AMR = ifelse(is.na(AMR), 0, AMR)) %>%
  rename(ID = deid_enc_id) %>%                     # Rename to ID for convenience
  relocate(HACDIF, .before = 2)                    # Relocate HACDIF as 2nd column

df_CDI_final <- df_CDI %>% 
  select(-c(age, admit_time, CDIF_order_time, CDIF_result_time, discharge_time))

# Old Code (keep for when join diabetes tbl)
#df_CDI <- df %>% 
 # left_join(meds_before, by = 'deid_enc_id') %>%   # Left join dm and med_taken_before_HACIDF
 # left_join(dbm, by = "deid_enc_id") %>%           # Left join dm and dbm
 # mutate(PPI = ifelse(is.na(PPI), 0, PPI),         # Reassign NA values for PPI:DBM to 0
   #      GAS = ifelse(is.na(GAS), 0, GAS),
   #      anyGAS = ifelse(is.na(anyGAS), 0, anyGAS),
   #      LAX = ifelse(is.na(LAX), 0, LAX),
   #      ABX = ifelse(is.na(ABX), 0, ABX),
   #      AMR = ifelse(is.na(ABX), 0, AMR),
   #      DBM = ifelse(is.na(DBM), 0, DBM)) %>%
 # rename(ID = deid_enc_id) %>%                     # Rename to ID for convenience
 # relocate(HACDIF, .before = 2)                    # Relocate HACDIF as 2nd column

# Change all to factor vars?
# df_CDI <- mutate_at(df_CDI, vars(c(readmit, age_gte_65, HACDIF, PPI, GAS, anyGAS, LAX, ABX, AMR)), factor)
```

# Preview final dataset

```{r Preview and explore `df_CDI_final`}
str(df_CDI_final)
summary(df_CDI_final)
```

# Calculate descriptives for categorical variables in `df_CDI_final`

```{r Create table one of descriptives}
tableone <- CreateCatTable(data=df_CDI_final, vars=c("readmit", "ED_dispo", "age_gte_65", "HACDIF", "PPI", "GAS", "anyGAS", "LAX", "ABX", "AMR"))
  # Edit later to include diabetes, DBM

print(tableone, showAllLevels = T, noSpaces = T)

# For binary variables, 1 = YES and 0 = NO. 
```

# Save resulting dataset `df_CDI_final` as CSV and RDS file

```{r Save resulting dataset as CSV and RDS}
write_csv(df_CDI_final, "/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/clean/df_CDI_risk_model.csv")
saveRDS(df_CDI_final, "/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/clean/df_CDI_risk_model.rds")
```
