---
title: "Prepare Dataset for CDI Risk Assessment Model"
author: "Steph Reynolds"
date: "`r format(Sys.time(), '%b %d, %Y  %H:%M %p')`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Background

## Project 
MINDSCAPE: Modeling of infectious network dynamics for surveillance, control and prevention enhancement

## Description 
This file reads in necessary data and identifies variables of interest for building a multivariate logistic regression model to evaluate C diff risk among hospitalized patients. The resulting dataframe (`df.CDI`) contains binary indicators for CDI risk factors. Variables include:

  * `ID` --> each patient's unique de-identified encounter ID 
  * `readmit` --> whether patient had a prior hospital admission in the last 30 days
  * `ED_dispo` --> whether pt was admitted from Emergency Dept (ED) ???
    - NOTE: UNCLEAR! VERIFY / LEO WHAT THIS SHOULD BE TO INDICATE THAT PT WAS ADMITTED TO ED!
  * `age_gte_65` --> whether pt is aged 65 years or older at time of hospitalization
  * `PPI` --> whether pt received proton pump inhibitor (PPI) during hospital stay
  * `GAS` --> whether pt received gastric acid suppressants and laxatives (GAS) during hospital stay
    - NOTE: This includes "ANTACIDS", "LAXATIVES AND CATHARTICS", and "LAXATIVES, LOCAL/RECTAL"
  * `ABX` --> whether pt received antibiotics during hospital stay 
    - NOTE: THIS DOES NOT INCLUDE OUTPATIENT ABX - ONLY THOSE DURING HOSPITALIZATION. 
  * `DBM` --> whether pt is diagnosed with diabetes mellitus (DBM) 
    - NOTE: DO NOT USE! ONLY HAVE DBM DX FOR COVID POSITIVE PTS.

# Load required packages 
```{r Load required packages}
library(readr)
library(tidyverse)
library(lubridate)
library(tableone)
library(kableExtra)
library(DT)
```

# Import `dm`, `med`, and `dx` tables 
```{r Import data}
dm <- read_delim("/Users/sreynolds2/Downloads/ucsf_data_pull_11.08.21/demographics and event table 11.08.21.csv", delim = "|")

med <- read_delim("/Users/sreynolds2/Downloads/ucsf_data_pull_11.08.21/med admin table 11.08.21.csv", delim = "|")

dx <- read_delim("/Users/sreynolds2/Downloads/ucsf_data_pull_11.08.21/covid diagnoses 11.08.21.csv", delim = "|")
```

# Clean `dm` table - deselect unnecessary variables, change to appropriate variable type
```{r Clean `dm` table}
dm <- dm %>% 
  select(deid_enc_id, age, readmit, HAI_type, HAI_first_date, ED_dispo, admit_time) %>% 
  mutate_at(c("readmit", "ED_dispo", "HAI_type"), factor) %>% 
  mutate(HAI_first_date = substr(HAI_first_date, 1,20)) %>% 
  mutate_at(c("admit_time", "HAI_first_date"), ymd_hms) %>% 
  mutate(age_gte_65 = if_else(age>=65, 1, 0),
         readmit = if_else(readmit=="Yes", 1, 0),
         cdif = if_else(HAI_type=="CDIF", 1, 0),
         HAI_gt_2_days = if_else((difftime(HAI_first_date, admit_time, units = "days")) > 2, 1, 0),
         # Add column to indicate whether CDIFF dx was > 2 days of admission
         HACDIF = if_else(cdif==1 & HAI_gt_2_days==1, 1, 0)) %>% 
  select(-c(age, HAI_type, admit_time, cdif, HAI_gt_2_days))

# Filter to obtain table of pts who were diagnosed with hospital-acquired C diff (HACDIF)
HA_cdif_pts <- dm %>% 
  filter(HACDIF==1)

# Proportion patients with HACDIF out of total number of pts in dm table
print(n_distinct(HA_cdif_pts$deid_enc_id)/(n_distinct(dm$deid_enc_id)))
```

# Clean `med` table - deselect unnecessary variables, change to appropriate variable type
```{r Clean `med` table}
# With med table, create new variables to indicate whether pt received PPI, GAS, and/or ABX
med <- med %>% 
  select(deid_enc_id, thera_class, pharm_class, taken_time) %>% 
  mutate_at(c("thera_class", "pharm_class"), factor) %>%
  mutate(PPI = if_else(pharm_class=="PROTON-PUMP INHIBITORS", 1, 0),    
         GAS = if_else(pharm_class %in% c("ANTACIDS", "LAXATIVES AND CATHARTICS", "LAXATIVES, LOCAL/RECTAL"), 1, 0),
         ABX = if_else(thera_class %in% c("ANTIBIOTICS", "ANTIFUNGALS", "ANTIINFECTIVES/MISCELLANEOUS", "ANTIPARASITICS", "ANTIVIRALS"), 1, 0)) %>% 
  select(-c(thera_class, pharm_class))
```

# Create `med_taken_before_HAI` table to indicate whether pt received PPI, GAS, or ABX up to one day before HAI_first_date (aka when HACDIF culture was ordered)
```{r Create `med_taken_before_HAI` table}
med_taken_before_HAI <- dm %>% 
 left_join(med, by = "deid_enc_id") %>%     # Left join dm and med tables on ID
  filter(taken_time < HAI_first_date) %>%     # Filter for where  taken_time (for PPI, ABX, or GAS) < HAI_first_date
  group_by(deid_enc_id) %>%                   # Group by ID
  summarize(Num_of_PPI = sum(PPI),            # Count total number of PPI, GAS, ABX up to day before HAI_first_date
            Num_of_GAS = sum(GAS),
            Num_of_ABX = sum(ABX)) %>% 
  mutate(PPI = if_else(Num_of_PPI>=1, 1, 0),  # Transform to binary vars, indicating yes/no whether pt received PPI, GAS, or ABX
         GAS = if_else(Num_of_GAS>=1, 1, 0),
         ABX = if_else(Num_of_ABX>=1, 1, 0)) %>% 
  select(deid_enc_id, PPI, GAS, ABX)          # Only keep relevant variables

# View table
med_taken_before_HAI
# Rather than number of PPI/GAS/ABX, should this just be a binary variable indicating receipt of PPI/GAS/ABX (coded as 1=yes or 0=no)? Section commented above handles this. Can remove this chunk of code if decide it's better to have frequency counts for PPI/GAS/ABX. 
```

# Clean `dx` table - create table `dbm` to indicate diabetes mellitus (DBM) diagnoses
```{r Clean `dx` table, create table to indicates DBM dx}
dbm <- dx %>% 
  mutate(DBM = if_else(dx_group=="DIABETES MELLITUS", 1, 0)) %>%
  group_by(deid_enc_id) %>% 
  summarize(n_dbm = sum(DBM)) %>% 
  mutate(DBM = if_else(n_dbm>=1, 1, 0)) %>% 
  select(-n_dbm)

sum(dbm$DBM)
# 406 patients diagnosed with diabetes 

# 406/1210 = 34% of Covid-positive pts diag with DBM 
```

NOTE: Only have diabetes diagnoses for 1,210 out of 54,955 patients. Missing data on DBM diagnosis for ~97% of patients. Only have DBM data for Covid-positive patients. Checked demographics and events table, and diabetes dx not listed as variable. Would need Leo to re-pull data gathering diabetes dx for ALL patients. 

# Join `dm`, `med`, and `dbm` tables and assign to `df_CDI`
```{r Merge all tables}
df_CDI <- dm %>% 
  left_join(med_taken_before_HAI, by = 'deid_enc_id') %>% # Left join dm and med_taken_before_HAI
  left_join(dbm, by = "deid_enc_id") %>%          # Left join dm and dbm
  mutate(PPI = ifelse(is.na(PPI), 0, PPI),        # Reassign NA values for PPI:DBM to 0
         GAS = ifelse(is.na(GAS), 0, GAS),
         ABX = ifelse(is.na(ABX), 0, ABX),
         DBM = ifelse(is.na(DBM), 0, DBM)) %>%
  rename(ID = deid_enc_id)                        # Rename to ID for convenience

df_CDI <- mutate_at(df_CDI, vars(c(readmit, age_gte_65, HACDIF, PPI, GAS, ABX, DBM)), factor)
  
```

# Preview and explore resulting dataset
```{r Preview and explore `df_CDI`}
# Preview dataset and check variables types 
str(df_CDI)

# Confirm that `df_CDI` has 55,037 obs - same number of obs as `dm` table

# Use DT package to explore dataset
DT::datatable(df_CDI)
```

# Calculate descriptives for categorical variables in `df_CDI`
```{r Create table one of descriptives}
tableone <- CreateCatTable(data=df_CDI, vars=c("readmit", "ED_dispo", "age_gte_65", "HACDIF", "PPI", "GAS", "ABX", "DBM"))

print(tableone, showAllLevels = T, noSpaces = T)

# For binary variables, 1 = YES and 0 = NO. 
```

# Save resulting dataset `df_CDI` as .csv file and rds object 
```{r Save resulting dataset as CSV and RDS}
write_csv(df_CDI, "/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/clean/df_CDI_risk_model.csv")
saveRDS(df_CDI, "/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/clean/df_CDI_risk_model.rds")
```

# Create 2x2 contingency table for age and CDIF and compute OR 
```{r Contingency table and OR for age and CDIF}
table(df_CDI$HACDIF, df_CDI$age_gte_65)

chisq.test(df_CDI$HACDIF, df_CDI$age_gte_65)
# p-value<0.05, thus, HACIF and AGE are NOT independent, aka they are related

fisher.test(df_CDI$HACDIF, df_CDI$age_gte_65)
# p-value<0.05, thus, HACIF are NOT independent, aka they are related
# OR=1.66, meaning patients that are 65 and older are 1.66x more likely to be diagnosed with HA-CDIF than those who are under 65 years old
```

# Create 2x2 contingency table for ABX and CDIF and compute OR 
```{r Contingency table and OR for ABX and CDIF}
table(df_CDI$HACDIF, df_CDI$ABX)

chisq.test(df_CDI$HACDIF, df_CDI$ABX)
# p-value<0.05, thus, HACIF and ABX are DEPENDENT

fisher.test(df_CDI$HACDIF, df_CDI$ABX)
# OR=633, meaning patients who received ABX up to one day before HAI diagnosis are 633x more likely to be diagnosed with HA-CDIF than those who do not receive ABX
```

# Create 2x2 contingency table for GAS and CDIF and compute OR 
```{r Contingency table and OR for GAS and CDIF}
table(df_CDI$HACDIF, df_CDI$GAS)

chisq.test(df_CDI$HACDIF, df_CDI$GAS)
# p-value<0.05, thus, HACIF and GAS are DEPENDENT

fisher.test(df_CDI$HACDIF, df_CDI$GAS)
# OR=633, meaning patients who received GAS up to one day before HAI diagnosis are 508x more likely to be diagnosed with HA-CDIF than those who do not receive GAS
```

# Create 2x2 contingency table for PPI and CDIF and compute OR 
```{r Contingency table and OR for PPI and CDIF}
table(df_CDI$HACDIF, df_CDI$PPI)

chisq.test(df_CDI$HACDIF, df_CDI$PPI)
# p<0.05, thus, HACIF and PPI are DEPENDENT

fisher.test(df_CDI$HACDIF, df_CDI$PPI)
# OR=1418, meaning patients who received a proton pump inhibitor (PPI) up to one day before HAI diagnosis are 141x more likely to be diagnosed with HA-CDIF than those who do not receive PPI
```

# Next Steps 

  * Create multivariate logistic regression model, divide data into 3 chunks/time periods (e.g. 2019, 2020, 2021)
    + 1 chunk used to train model, the other 2 used to evaluate model performance (e.g. 2019 chunk used to train model, while 2020-2021 used to test model performance)
  * Evaluate "model creep" or how much the optimized model changes over time? 
