---
title: "Prepare Dataset for CDI Risk Assessment Model - Edits (Combine PPI and gastric acid sup)"
author: "Steph Reynolds"
date: "`r format(Sys.time(), '%b %d, %Y  %H:%M %p')`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Background

## Project

MINDSCAPE: Modeling of infectious network dynamics for surveillance, control and prevention enhancement

## Description

This file reads in necessary data and identifies variables of interest for building a multivariate logistic regression model to evaluate C diff risk among hospitalized patients. The resulting data frame (`df.CDI`) contains binary indicators for CDI risk factors. Variables include:

-   `ID` --\> each patient's unique de-identified encounter ID
-   `readmit` --\> whether patient had a prior hospital admission in the last 30 days
-   `ED_dispo` --\> whether pt was admitted from Emergency Dept (ED) ???
    -   NOTE: UNCLEAR! VERIFY / LEO WHAT THIS SHOULD BE TO INDICATE THAT PT WAS ADMITTED TO ED!
-   `age_gte_65` --\> whether pt is aged 65 years or older at time of hospitalization
-   `PPI` --\> whether pt received proton pump inhibitor (PPI) during hospital stay
-   `GAS` --\> whether pt received gastric acid suppressants (GAS) during hospital stay
    -   NOTE: This includes "ANTACIDS" and "HISTAMINE H2-RECEPTOR INHIBITORS"
-   `anyGAS` --\> whether pt received PPI and/or GAS during hospital stay
-   `LAX` --\> whether pt received laxatives (LAX) during hospital stay
    -   NOTE: This includes "LAXATIVES AND CATHARTICS" and "LAXATIVES, LOCAL/RECTAL"
-   `ABX` --\> whether pt received antibiotics during hospital stay
    -   NOTE: THIS DOES NOT INCLUDE OUTPATIENT ABX - ONLY THOSE DURING HOSPITALIZATION.
-   `DBM` --\> whether pt is diagnosed with diabetes mellitus (DBM)
    -   NOTE: DO NOT USE! ONLY HAVE DBM DX FOR COVID POSITIVE PTS.

# Load required packages

```{r Load required packages}
library(readr)
library(tidyverse)
library(lubridate)
library(tableone)
library(DescTools) # need for %like% 
```

# Import `dm`, `med`, and `dx` tables

```{r Import data}
dm <- read_delim("/Users/sreynolds2/Downloads/ucsf_data_pull_11.08.21/demographics and event table 11.08.21.csv", delim = "|")

med <- read_delim("/Users/sreynolds2/Downloads/ucsf_data_pull_11.08.21/med admin table 11.08.21.csv", delim = "|")

dx <- read_delim("/Users/sreynolds2/Downloads/ucsf_data_pull_11.08.21/covid diagnoses 11.08.21.csv", delim = "|")
```

# Clean `dm` table: Deselect unnecessary variables and change to appropriate variable type

```{r Clean `dm` table}
dm <- dm %>% 
  select(deid_enc_id, age, readmit, HAI_type, HAI_first_date, ED_dispo, admit_time) %>% 
  mutate_at(c("readmit", "ED_dispo", "HAI_type"), factor) %>% 
  mutate(HAI_first_date = substr(HAI_first_date, 1, 20)) %>% 
  mutate_at(c("admit_time", "HAI_first_date"), ymd_hms) %>% 
  mutate(age_gte_65 = if_else(age>=65, 1, 0),
         readmit = if_else(readmit=="Yes", 1, 0),
         cdif = if_else(HAI_type=="CDIF", 1, 0),
         HAI_gt_2_days = if_else((difftime(HAI_first_date, admit_time, units = "days")) > 2, 1, 0),
         # Add column to indicate whether CDIFF dx was > 2 days of admission
         HACDIF = if_else(cdif==1 & HAI_gt_2_days==1, 1, 0)) %>% 
  select(-c(age, HAI_type, admit_time, cdif, HAI_gt_2_days))

# Filter for pts who were diagnosed with hospital-acquired C diff (HACDIF)
HA_cdif_pts <- dm %>% 
  filter(HACDIF==1)

# Proportion patients with HACDIF out of total number of pts in dm table
cat("Proportion of pts with HA-CDI out of total number of pts:", 
    round(n_distinct(HA_cdif_pts$deid_enc_id)/(n_distinct(dm$deid_enc_id)), 4))
```

# Create binary variables to indicate whether patients received PPI, GAS, LAX, and/or ABX; assign to table `med2`

**Alternative method - didn't work when I re-ran on 3/7/22, but keep just in case**

```{r, eval=FALSE}
# With med table, create new variables to indicate whether pt received PPI, GAS, anyGAS, LAX and ABX
# See beginning of script for description of variables 

med2 <- med %>% 
  select(deid_enc_id, name, thera_class, pharm_class, taken_time) %>% 
  mutate_at(c("thera_class", "pharm_class"), factor) %>%
  mutate(PPI = if_else(pharm_class=="PROTON-PUMP INHIBITORS" | str_detect(name, "PANTOPRAZOLE"), 1, 0),    
         GAS = if_else(pharm_class %in% c("ANTACIDS", "HISTAMINE H2-RECEPTOR INHIBITORS"), 1, 0),
         anyGAS = if_else(pharm_class %in% c("PROTON-PUMP INHIBITORS", "ANTACIDS", "HISTAMINE H2-RECEPTOR INHIBITORS") | str_which(name, "PANTOPRAZOLE"), 1, 0),
         LAX = if_else(pharm_class %in% c("LAXATIVES AND CATHARTICS", "LAXATIVES, LOCAL/RECTAL"), 1, 0),
         ABX = if_else(thera_class %in% c("ANTIBIOTICS", "ANTIFUNGALS", "ANTIINFECTIVES/MISCELLANEOUS", "ANTIPARASITICS","ANTIVIRALS") | str_which(name, "VANCOMYCIN|CEFTRIAXONE|CEFEPIME|LEVOFLOXACIN|CIPROFLOXACIN"), 1, 0))
```

# Create binary variables to indicate whether patients received PPI, GAS, LAX, and/or ABX; assign to table `med2`

```{r Create binary vars to indicate whether pts received PPI, GAS, LAX, ABX}
# Create vector for names of common antibiotics
names_of_abx <- "VANCOMYCIN|CEFTRIAXONE|CEFEPIME|LEVOFLOXACIN|CIPROFLOXACIN|AMIKACIN|GENTAMICIN|NEOMYCIN-POLYMYXIN|TOBRAMYCIN|ERTAPENEM|IMIPENEM-CILASTATIN|MEROPENEM|MEROPENEM-VABORBACTAM|CEFAZOLIN|CEFOXITIN|CEFOTETAN|CEFUROXIME|CEFTAZIDIME|CEFTAROLINE|TOBRAMYCIN|DALBAVANCIN|ORITAVANCIN|CLINDAMYCIN|DAPTOMYCIN|AZITHROMYCIN|ERYTHROMYCIN|AZTREONAM|LINEZOLID|AMPICILLIN|AMPICILLIN-SULBACTAM|NAFCILLIN|PENICILLIN|PIPERACILLIN-TAZOBACTAM|BACITRACIN|NEOMYCIN-POLYMYXIN|GENTAMICIN-POLYMYXIN|TRIMETHOPRIM-SULFAMETHOXAZOLE|SULFAMETHOXAZOLE-TRIMETHOPRIM|DOXYCYCLINE|MINOCYCLINE|CLOFAZIMINE|ISONIAZID|RIFAMPIN|TIGECYCLINE|TRIMETHOPRIM|ACYCLOVIR|CASPOFUNGIN|LETERMOVIR|GANCICLOVIR|ACYCLOVIR|POSACONAZOLE|AMPHOTERICIN|AZTREONAM|GENTAMICIN|CASIRIVIMAB|CEFIDEROCOL|CEFTOLOZANE-TAZOBACTAM|CIDOFOVIR|CLINDAMYCIN|COLISTIMETHATE|ALBAVANCIN|ERAVACYCLINE|FOSCARNET|GENTAMICIN-POLYMYXIN|REMDESIVIR|ISAVUCONAZONIUM|LINEZOLID|MASTOID POWDER|MICAFUNGIN|CEFAZOLIN/CEFTAZIDIME|ORITAVANCIN|PENICILLIN|PERAMIVIR|TEDIZOID|TRIMETHOPRIM-SULFAMETHOXAZOLE|VORICONAZOLE|BACITRACIN|IBALIZUMAB-UIYK|CENICRIVIROC|BAMLANIVIMAB-ETESEVIMAB|BAMLANIVIMAB|BEZLOTOXUMAB"

# Create variables to indicate whether pt received PPI, GAS, anyGAS, LAX and ABX
med2 <- med %>% 
  select(deid_enc_id, name, thera_class, pharm_class, taken_time) %>% 
  mutate_at(c("thera_class", "pharm_class"), factor) %>%
  mutate(PPI = if_else(pharm_class=="PROTON-PUMP INHIBITORS" | name %like% "PANTOPRAZOLE", 1, 0),    
         GAS = if_else(pharm_class %in% c("ANTACIDS", "HISTAMINE H2-RECEPTOR INHIBITORS"), 1, 0),
         anyGAS = if_else(pharm_class %in% c("PROTON-PUMP INHIBITORS", "ANTACIDS", "HISTAMINE H2-RECEPTOR INHIBITORS") | name %like% "PANTOPRAZOLE", 1, 0),
         LAX = if_else(pharm_class %in% c("LAXATIVES AND CATHARTICS", "LAXATIVES, LOCAL/RECTAL"), 1, 0),
         ABX = if_else(thera_class %in% c("ANTIBIOTICS", "ANTIFUNGALS", "ANTIINFECTIVES/MISCELLANEOUS", "ANTIPARASITICS","ANTIVIRALS") | name %like% names_of_abx, 1, 0))
```

# Create `med_taken_before_HAI` table to indicate whether pt received PPI, GAS, or ABX up to one day before HAI_first_date\* \*

\*`HAI_first_date` is when HACDIF culture was ordered

```{r Create `med_taken_before_HAI` table}

full <- dm %>% 
  left_join(med2, by = "deid_enc_id") %>% 
  filter(ABX==0 & thera_class=="NULL") %>% 
  select(deid_enc_id, name, thera_class, pharm_class)

full %>% distinct(name, .keep_all = T) %>% write_csv("/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/check_for_abx.csv")


####
full <- full %>% 
  mutate(abx_yn = case_when(ABX==1 & (taken_time < HAI_first_date | is.na(HAI_first_date)) ~ 1))

full %>% summarize(n_distinct(deid_enc_id)) 



med_taken_before_HAI <- dm %>% 
  left_join(med2, by = "deid_enc_id") %>%     # Left join dm and med tables on ID
  filter(taken_time < HAI_first_date) %>%     # Filter for where  taken_time (for PPI, ABX, or GAS) < HAI_first_date
  group_by(deid_enc_id) %>%                   # Group by ID
  summarize(Num_of_PPI = sum(PPI),            # Count total number of PPI, GAS, anyGAS, LAX, ABX up to day before HAI_first_date
            Num_of_GAS = sum(GAS),
            Num_of_anyGAS = sum(anyGAS),
            Num_of_LAX = sum(LAX),
            Num_of_ABX = sum(ABX)) %>% 
  mutate(PPI = if_else(Num_of_PPI>=1, 1, 0),  # Transform to binary vars, indicating yes/no whether pt received PPI, GAS, or ABX
         GAS = if_else(Num_of_GAS>=1, 1, 0),
         anyGAS = if_else(Num_of_anyGAS>=1, 1, 0),
         LAX = if_else(Num_of_LAX>=1, 1, 0),
         ABX = if_else(Num_of_ABX>=1, 1, 0)) %>% 
  select(deid_enc_id, PPI, GAS, anyGAS, LAX, ABX)          # Only keep relevant variables

# View table
med_taken_before_HAI

# Rather than number of PPI/GAS/anyGAS/LAX/ABX, should this just be a binary variable indicating receipt of PPI/GAS/anyGAS/LAX/ABX (coded as 1=yes or 0=no)? Section commented above handles this. Can remove this chunk of code if decide it's better to have frequency counts for PPI/GAS/anyGAS/LAX/ABX. 


# REPLICATE CODE BUT KEEP FREQUENCY COUNTS OF ABX, INSTEAD OF CONVERTING TO BINARY VARIABLE (Y/N).
med_taken_before_HAI_2 <- dm %>% 
  left_join(med2, by = "deid_enc_id") %>%     # Left join dm and med tables on ID
  filter(taken_time < HAI_first_date) %>%     # Filter for where  taken_time (for PPI, ABX, or GAS) < HAI_first_date
  group_by(deid_enc_id) %>%                   # Group by ID
  summarize(Num_of_PPI = sum(PPI),            # Count total number of PPI, GAS, anyGAS, LAX, ABX up to day before HAI_first_date
            Num_of_GAS = sum(GAS),
            Num_of_anyGAS = sum(anyGAS),
            Num_of_LAX = sum(LAX),
            Num_of_ABX = sum(ABX)) 

num_of_abx <- med_taken_before_HAI_2 %>% select(c(deid_enc_id, Num_of_ABX))
num_of_abx %>% filter(Num_of_ABX!=0)
sum(med_taken_before_HAI_2$Num_of_ABX)
```

# Clean `dx` table and create binary variable to indicate whether patient has diabetes; assign to table `dx`

```{r Clean `dx` table, create table to indicates DBM dx}
dbm <- dx %>% 
  mutate(DBM = if_else(dx_group=="DIABETES MELLITUS", 1, 0)) %>%
  group_by(deid_enc_id) %>% 
  summarize(n_dbm = sum(DBM)) %>% 
  mutate(DBM = if_else(n_dbm>=1, 1, 0)) %>% 
  select(-n_dbm)

sum(dbm$DBM)
# 406 patients diagnosed with diabetes 

# 406/1210 = 34% of Covid-positive pts diag with DBM 
```

**NOTE:** Only have diabetes diagnoses for 1,210 out of 54,955 patients. Missing data on DBM diagnosis for \~97% of patients. Only have DBM data for Covid-positive patients. Checked demographics and events table, and diabetes dx not listed as variable. Would need Leo to re-pull data gathering diabetes dx for ALL patients.

# Join `dm`, `med`, and `dbm` tables and assign to `df_CDI`

```{r Merge all tables}
df_CDI <- dm %>% 
  left_join(med_taken_before_HAI, by = 'deid_enc_id') %>% # Left join dm and med_taken_before_HAI
  left_join(dbm, by = "deid_enc_id") %>%          # Left join dm and dbm
  mutate(PPI = ifelse(is.na(PPI), 0, PPI),        # Reassign NA values for PPI:DBM to 0
         GAS = ifelse(is.na(GAS), 0, GAS),
         anyGAS = ifelse(is.na(anyGAS), 0, anyGAS),
         LAX = ifelse(is.na(LAX), 0, LAX),
         ABX = ifelse(is.na(ABX), 0, ABX), 
         DBM = ifelse(is.na(DBM), 0, DBM)) %>%
  rename(ID = deid_enc_id)                        # Rename to ID for convenience

df_CDI <- mutate_at(df_CDI, vars(c(readmit, age_gte_65, HACDIF, PPI, GAS, anyGAS, LAX, ABX, DBM)), factor)
```

# View resulting dataset

```{r Preview and explore `df_CDI`}
# Preview dataset and check variables types 
str(df_CDI)
glimpse(df_CDI)

# Confirm that `df_CDI` has 55,037 obs - same number of obs as `dm` table

summary(df_CDI)
```

# Calculate descriptives for categorical variables in `df_CDI`

```{r Create table one of descriptives}
tableone <- CreateCatTable(data=df_CDI, vars=c("readmit", "ED_dispo", "age_gte_65", "HACDIF", "PPI", "GAS", "anyGAS", "LAX", "ABX", "DBM"))

print(tableone, showAllLevels = T, noSpaces = T)

# For binary variables, 1 = YES and 0 = NO. 
```

# Save resulting dataset `df_CDI` as CSV and RDS file

```{r Save resulting dataset as CSV and RDS}
write_csv(df_CDI, "/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/clean/df_CDI_risk_model.csv")
saveRDS(df_CDI, "/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/clean/df_CDI_risk_model.rds")
```

# Create 2x2 contingency table for age and CDIF and compute OR

```{r Contingency table and OR for age and CDIF}
table(df_CDI$HACDIF, df_CDI$age_gte_65)

chisq.test(df_CDI$HACDIF, df_CDI$age_gte_65)

fisher.test(df_CDI$HACDIF, df_CDI$age_gte_65)
```

**Results:**

p\<0.05, thus, HACIF and AGE are NOT independent, aka they are related

OR=1.66, meaning patients that are 65 and older are 1.66x more likely to be diagnosed with HA-CDIF than those who are under 65 years old

# Create 2x2 contingency table for ABX and CDIF and compute OR

```{r Contingency table and OR for ABX and CDIF}
table(df_CDI$HACDIF, df_CDI$ABX)

chisq.test(df_CDI$HACDIF, df_CDI$ABX)

fisher.test(df_CDI$HACDIF, df_CDI$ABX)
```

**Results:**

p\<0.05, thus, HACIF and ABX are DEPENDENT

OR=963, meaning patients who received ABX up to one day before HAI diagnosis are 963x more likely to be diagnosed with HA-CDIF than those who do not receive ABX

# Create 2x2 contingency table for GAS and CDIF and compute OR

```{r Contingency table and OR for GAS and CDIF}
table(df_CDI$HACDIF, df_CDI$GAS)

chisq.test(df_CDI$HACDIF, df_CDI$GAS)

fisher.test(df_CDI$HACDIF, df_CDI$GAS)
```

**Results:**

p\<0.05, thus, HACIF and GAS are DEPENDENT

OR=233, meaning patients who received gastric acid suppressants up to one day before HAI diagnosis are 233x more likely to be diagnosed with HA-CDIF than those who do not receive gastric acid suppressants

# Create 2x2 contingency table for anyGAS and CDIF and compute OR

```{r Contingency table and OR for anyGAS and CDIF}
table(df_CDI$HACDIF, df_CDI$anyGAS)

chisq.test(df_CDI$HACDIF, df_CDI$anyGAS)

fisher.test(df_CDI$HACDIF, df_CDI$anyGAS)
```

**Results:**

p\<0.05, thus, HACIF and anyGAS are DEPENDENT

OR=343, meaning patients who received gastric acid suppressants and/or PPI up to one day before HAI diagnosis are 343x more likely to be diagnosed with HA-CDIF than those who do not receive gastric acid suppressants and/or PPI

# Create 2x2 contingency table for LAX and CDIF and compute OR

```{r Contingency table and OR for LAX and CDIF}
table(df_CDI$HACDIF, df_CDI$LAX)

chisq.test(df_CDI$HACDIF, df_CDI$LAX)

fisher.test(df_CDI$HACDIF, df_CDI$LAX)
```

**Results:**

p\<0.05, thus, HACIF and LAX are DEPENDENT

OR=409, meaning patients who received laxatives up to one day before HAI diagnosis are 409x more likely to be diagnosed with HA-CDIF than those who do not receive laxatives

# Create 2x2 contingency table for PPI and CDIF and compute OR

```{r Contingency table and OR for PPI and CDIF}
table(df_CDI$HACDIF, df_CDI$PPI)

chisq.test(df_CDI$HACDIF, df_CDI$PPI)

fisher.test(df_CDI$HACDIF, df_CDI$PPI)
```

**Results:**

p\<0.05, thus, HACIF and PPI are DEPENDENT

OR=141, meaning patients who received a proton pump inhibitor (PPI) up to one day before HAI diagnosis are 141x more likely to be diagnosed with HA-CDIF than those who do not receive PPI
