---
title: "Univariate and Multivariate Regression Analysis Summary Tables for Hospital Acquired CDI Risk"
author: "Steph Reynolds"
date: "`r format(Sys.time(), '%b %d, %Y  %H:%M %p')`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Background

## Project

MINDSCAPE: Modeling of infectious network dynamics for surveillance, control and prevention enhancement

## Description

This file reads in `df_CDI_risk_model.csv` and displays univariate and multivariate regression model results in formatted, publication-ready tables. Both models aim to predict risk of hospital-acquired *Clostridioides difficile* infection (HA-CDI).

# Load required packages

```{r Load required packages}
library(tidyverse)
library(gtsummary)
library(performance)
```

# Import HA-CDI dataset and transform variables to factor type

```{r Import data}
df <- read_csv("/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/clean/df_CDI_risk_model.csv") %>% 
  select(-ID) %>% 
  mutate_if(is.numeric, factor)
```

# Reassign levels for categorical variables

```{r Convert to proper variable types }
# Reassign levels to No and Yes for factor variables
for (i in colnames(df)){
   levels(df[[i]]) <- c("No", "Yes")
}

# Check structure and var types
str(df)
```

# Create table of descriptive statistics

```{r Table of summary statistics}
# Set theme for gtsummary 
theme_gtsummary_journal(journal = "jama")
theme_gtsummary_compact()

# Table of descriptive stats 
df %>% 
  tbl_summary(by = HACDIF, 
              statistic = all_categorical() ~ "{n} / {N} ({p}%)",
              label = list(readmit ~ "Readmission in last 30 days",
                           age_gte_65 ~ "Age>=65 years old",
                           ED ~ "Transfer from Emergency Department",
                           PPI ~ "Receipt of proton pump inhibitor (PPI)",
                           GAS ~ "Receipt of antacids or H2R antagonist (GAS)",
                           anyGAS ~ "Receipt of GAS or PPI",
                           LAX ~ "Receipt of laxative",
                           ABX ~ "Receipt of antibiotic",
                           AMR ~ "Receipt of antimicrobial")) %>% 
  add_p() %>% 
  add_overall() %>%
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Positive hospital-acquired C difficile infection**") %>%
  modify_footnote(all_stat_cols() ~ "N (%)") %>% 
  modify_caption("**Table 1. Patient Characteristics**") %>% 
  as_gt()

```

# Create summary table of univariate regression model results

```{r Summary table of univariate regression}
tbl_uv_reg <- df %>%
    tbl_uvregression(
    method = glm, 
    y = HACDIF,
    method.args = list(family = binomial),
    exponentiate = T, 
    pvalue_fun = ~style_pvalue(.x, digits = 2),
    label = list(readmit ~ "Readmission in last 30 days",
                 age_gte_65 ~ "Age>=65 years old",
                 ED ~ "Transfer from Emergency Department",
                 PPI ~ "Receipt of proton pump inhibitor (PPI)",
                 GAS ~ "Receipt of antacids or H2R antagonist (GAS)",
                 anyGAS ~ "Receipt of GAS or PPI",
                 LAX ~ "Receipt of laxative",
                 ABX ~ "Receipt of antibiotic",
                 AMR ~ "Receipt of antimicrobial"),
    show_single_row = c(readmit, age_gte_65, ED, PPI, GAS, anyGAS, LAX, ABX, AMR)) %>%
  modify_header(list(label ~ "**Variable**"), p.value = "**P**") %>%
  bold_p() %>% 
  modify_caption("**Table 2. Results of Univariate Regression**")

as_gt(tbl_uv_reg)
```

# Build and compare performance of multivariate logistic regression models

```{r Build multivariate log reg model}
# FULL MODEL includes ALL predictors 
m.full <- glm(data = df, formula = HACDIF ~ readmit + age_gte_65 + ED + PPI + GAS + anyGAS + LAX + ABX + AMR, family = binomial)
summary(m.full)

  # ED is not a signficant predictor, so we can remove it from our model. 

# Model 1 excludes ED, PPI, GAS, ABX
m1 <- glm(data = df, formula = HACDIF ~ readmit + age_gte_65 + anyGAS + LAX + AMR, family = binomial)
  # Summarize model output
  summary(m1)
  # Plot ROC 
  logit_prob <- predict(m1, newdata = df, type = 'response')
  logit_roc.m1 <- pROC::roc(df$HACDIF ~ logit_prob, plot = T, print.auc = T)

# Model 2 excludes ED, anyGAS, AMR
m2 <- glm(data = df, formula = HACDIF ~ readmit + age_gte_65 + PPI + GAS + LAX + ABX, family = binomial)
  # Summarize model output
  summary(m2)
  # Plot ROC 
  logit_prob <- predict(m2, newdata = df, type = 'response')
  logit_roc.m2 <- pROC::roc(df$HACDIF ~ logit_prob, plot = T, print.auc = T)

# Model 3 excludes ED, anyGAS, ABX
m3 <- glm(data = df, formula = HACDIF ~ readmit + age_gte_65 + PPI + GAS + LAX + AMR, family = binomial)
  # Summarize model output
  summary(m3)
  # Plot ROC 
  logit_prob <- predict(m3, newdata = df, type = 'response')
  logit_roc.m3 <- pROC::roc(df$HACDIF ~ logit_prob, plot = T, print.auc = T)

# Model 4 excludes ED, PPI, GAS, AMR
m4 <- glm(data = df, formula = HACDIF ~ readmit + age_gte_65 + anyGAS + LAX + ABX, family = binomial)
  # Summarize model output
  summary(m4)
  # Plot ROC 
  logit_prob <- predict(m4, newdata = df, type = 'response')
  logit_roc.m4 <- pROC::roc(df$HACDIF ~ logit_prob, plot = T, print.auc = T)
  
# Model 5 excludes ED, anyGAS, LAX, AMR 
m5 <- glm(data = df, formula = HACDIF ~ readmit + age_gte_65 + PPI + GAS + ABX, family = binomial)
  # Summarize model output
  summary(m5)
  # Plot ROC 
  logit_prob <- predict(m5, newdata = df, type = 'response')
  logit_roc.m5 <- pROC::roc(df$HACDIF ~ logit_prob, plot = T, print.auc = T)
  
# Compare performance for all models (calculate AIC, BIC, Tjur's R2, RMSE, Log loss)
    # Set rank=T to order by highest performing model
m.performance <- compare_performance(m.full, m1, m2, m3, m4, m5, rank = T, metrics = "common")
m.performance
```

### Summary

-   Since inherently `anyGAS` includes `GAS` and `PPI`, and `AMR` includes `ABX`, the full model has an issue with multicollinearity. This is illustrated by the high variance inflation factor (VIF) for `AMR` and `ABX`.

-   When comparing all models, since model 5 has the lowest AIC/BIC but highest AUROC and adjusted R\^2, we can assume that it is the best performing.

-   Model 5 includes the following predictors: `readmit`, `age_gte_65`, `PPI`, `GAS`, and `ABX`. We removed `ED` and `LAX` since they were not significant (p\>.05), as well as `anyGAS` and `AMR` due to multicollinearity as mentioned above.

#### *Performance Metrics of Model 5*

| Metric    | Value  |
|-----------|--------|
| AIC       | 3564.2 |
| Tjur's R2 | 0.010  |
| RMSE      | 0.077  |
| AUROC     | 0.817  |

# Create summary table of multivariate regression model results

```{r Summary table of multivariate regression}
tbl_mv_reg <- tbl_regression(m2, 
                             exponentiate = T,
                             label = list(readmit ~ "Readmission in last 30 days",
                                          age_gte_65 ~ "Age>=65 years old",
                                          PPI ~ "Receipt of proton pump inhibitor (PPI)",
                                          GAS ~ "Receipt of antacids or H2R antagonist (GAS)",
                                          LAX ~ "Receipt of laxative",
                                          ABX ~ "Receipt of antibiotic"), 
                             show_single_row = c(readmit, age_gte_65, PPI, GAS, LAX, ABX)) %>%
  modify_header(list(label ~ "**Variable**"), p.value = "**P**") %>%
  bold_p() %>% 
  modify_caption("**Table 3. Results of Multivariate Regression**")

as_gt(tbl_mv_reg)
```
