---
title: "Odds Ratios for HA-CDI Risk Factors"
author: "Steph Reynolds"
date: "`r format(Sys.time(), '%b %d, %Y  %H:%M %p')`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Background

## Project

MINDSCAPE: Modeling of infectious network dynamics for surveillance, control and prevention enhancement

## Description

This file reads in `df.CDI which contains binary variables or risk factors associated with hospital-acquired C difficile (HA-CDIF) and computes the odds ratios for each risk factor. Variables include:

-   `ID` --\> each patient's unique de-identified encounter ID
-   `readmit` --\> whether patient had a prior hospital admission in the last 30 days
-   `ED_dispo` --\> whether pt was admitted from Emergency Dept (ED) ???
    -   NOTE: UNCLEAR! VERIFY / LEO WHAT THIS SHOULD BE TO INDICATE THAT PT WAS ADMITTED TO ED!
-   `age_gte_65` --\> whether pt is aged 65 years or older at time of hospitalization
-   `PPI` --\> whether pt received proton pump inhibitor (PPI) during hospital stay
-   `GAS` --\> whether pt received gastric acid suppressants (GAS) during hospital stay
    -   NOTE: This includes "ANTACIDS" and "HISTAMINE H2-RECEPTOR INHIBITORS"
-   `anyGAS` --\> whether pt received PPI and/or GAS during hospital stay
-   `LAX` --\> whether pt received laxatives (LAX) during hospital stay
    -   NOTE: This includes "LAXATIVES AND CATHARTICS" and "LAXATIVES, LOCAL/RECTAL"
-   `ABX` --\> whether pt received antibiotics during hospital stay
    -   NOTE: THIS DOES NOT INCLUDE OUTPATIENT ABX - ONLY THOSE DURING HOSPITALIZATION.
-   `DBM` --\> whether pt is diagnosed with diabetes mellitus (DBM)
    -   NOTE: DO NOT USE! ONLY HAVE DBM DX FOR COVID-POSITIVE PTS.

# Load required packages

```{r Load required packages}
library(tidyverse)
```

# Read in `df_CDI`
```{r Read in data}
df_CDI <- readRDS("/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/clean/df_CDI_risk_model.rds")
```

# Create 2x2 contingency table for age and CDIF and compute OR

```{r Contingency table and OR for age and CDIF}
table(df_CDI$HACDIF, df_CDI$age_gte_65)

chisq.test(df_CDI$HACDIF, df_CDI$age_gte_65)

fisher.test(df_CDI$HACDIF, df_CDI$age_gte_65)
```

**Results:**

p\<0.05, thus, HACIF and AGE are NOT independent, aka they are related

OR=1.64, meaning patients that are 65 and older are 1.64x more likely to be diagnosed with HA-CDIF than those who are under 65 years old

# Create 2x2 contingency table for ABX and CDIF and compute OR

```{r Contingency table and OR for ABX and CDIF}
table(df_CDI$HACDIF, df_CDI$ABX)

chisq.test(df_CDI$HACDIF, df_CDI$ABX)

fisher.test(df_CDI$HACDIF, df_CDI$ABX)
```

**Results:**

p\<0.05, thus, HACIF and ABX are DEPENDENT

OR=34.9, meaning patients who received ABX up to one day before HAI diagnosis are 34.9x more likely to be diagnosed with HA-CDIF than those who do not receive ABX

# Create 2x2 contingency table for AMR and CDIF and compute OR

```{r Contingency table and OR for AMR and CDIF}
table(df_CDI$HACDIF, df_CDI$AMR)

chisq.test(df_CDI$HACDIF, df_CDI$AMR)

fisher.test(df_CDI$HACDIF, df_CDI$AMR)
```

**Results:**

p\<0.05, thus, HACIF and ABX are DEPENDENT

OR=29.1, meaning patients who received AMR up to one day before HAI diagnosis are 29.1 more likely to be diagnosed with HA-CDIF than those who do not receive AMR

# Create 2x2 contingency table for GAS and CDIF and compute OR

```{r Contingency table and OR for GAS and CDIF}
table(df_CDI$HACDIF, df_CDI$GAS)

chisq.test(df_CDI$HACDIF, df_CDI$GAS)

fisher.test(df_CDI$HACDIF, df_CDI$GAS)
```

**Results:**

p\<0.05, thus, HACIF and GAS are DEPENDENT

OR=3.71, meaning patients who received gastric acid suppressants up to one day before HAI diagnosis are 3.71x more likely to be diagnosed with HA-CDIF than those who do not receive gastric acid suppressants

# Create 2x2 contingency table for anyGAS and CDIF and compute OR

```{r Contingency table and OR for anyGAS and CDIF}
table(df_CDI$HACDIF, df_CDI$anyGAS)

chisq.test(df_CDI$HACDIF, df_CDI$anyGAS)

fisher.test(df_CDI$HACDIF, df_CDI$anyGAS)
```

**Results:**

p\<0.05, thus, HACIF and anyGAS are DEPENDENT

OR=4.44, meaning patients who received gastric acid suppressants and/or PPI up to one day before HAI diagnosis are 4.44x more likely to be diagnosed with HA-CDIF than those who do not receive gastric acid suppressants and/or PPI

# Create 2x2 contingency table for LAX and CDIF and compute OR

```{r Contingency table and OR for LAX and CDIF}
table(df_CDI$HACDIF, df_CDI$LAX)

chisq.test(df_CDI$HACDIF, df_CDI$LAX)

fisher.test(df_CDI$HACDIF, df_CDI$LAX)
```

**Results:**

p\<0.05, thus, HACIF and LAX are DEPENDENT

OR=2.01, meaning patients who received laxatives up to one day before HAI diagnosis are 2.01 more likely to be diagnosed with HA-CDIF than those who do not receive laxatives

# Create 2x2 contingency table for PPI and CDIF and compute OR

```{r Contingency table and OR for PPI and CDIF}
table(df_CDI$HACDIF, df_CDI$PPI)

chisq.test(df_CDI$HACDIF, df_CDI$PPI)

fisher.test(df_CDI$HACDIF, df_CDI$PPI)
```

**Results:**

p\<0.05, thus, HACIF and PPI are DEPENDENT

OR=3.11, meaning patients who received a proton pump inhibitor (PPI) up to one day before HAI diagnosis are 3.11 more likely to be diagnosed with HA-CDIF than those who do not receive PPI

# Create 2x2 contingency table for readmit and CDIF and compute OR

```{r Contingency table and OR for readmit and CDIF}
table(df_CDI$HACDIF, df_CDI$readmit)

chisq.test(df_CDI$HACDIF, df_CDI$readmit)

fisher.test(df_CDI$HACDIF, df_CDI$readmit)
```

**Results:**

p\<0.05, thus, HACIF and PPI are DEPENDENT

OR=1.92, meaning patients who were readmitted are 1.92 more likely to be diagnosed with HA-CDIF than those who were not 

