---
title: "Table One for CDI Risk Model"
author: "Steph Reynolds (Stephanie.Reynolds@ucsf.edu)"
date: "`r format(Sys.time(), '%b %d, %Y  %H:%M %p')`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Background

## Project

MINDSCAPE: Modeling of infectious network dynamics for surveillance, control and prevention enhancement

## Description

This scripts reads in the CDI risk factor dataset (`df_CDI_risk_model.csv`) and creates a Table 1 showing frequencies for all predictors stratified by hospital-acquired C difficile diagnosis.

## References

-   `table1`: Tables of Descriptive Statistics in HTML ([Link](https://cran.r-project.org/web/packages/table1/))

    -   Vignette: Using the table1 Package to Create HTML Tables of Descriptive Statistics (Rich, 2021)

# Install and load required packages

```{r Load required packages, include=F}
#install.packages("table1")
library(table1)
library(tidyverse)
```

# Import and preview data

```{r Read in df}
# Read in data and assign to `df`
df <- read_csv("/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/clean/df_CDI_risk_model.csv") %>% 
  select(-c(ID, ED_dispo)) %>%     # Exclude ED_dispo and age (continuous) for now 
  mutate_if(is.numeric, factor)

# Preview dataset 
glimpse(df)
```

# Reassign levels for categorical variables

```{r Convert to proper variable types }
# Reassign levels to No and Yes for factor variables
for (i in colnames(df)){
   levels(df[[i]]) <- c("No", "Yes")
}

# Check structure and var types
str(df)
```

# Assign labels to categorical variables

```{r Assign variable labels}
label(df$HACDIF) <- "Hospital-acquired C difficile infection"
label(df$readmit) <- "Readmission"
label(df$age_gte_65) <- "Age >= 65 years old"
label(df$PPI) <- "Receipt of proton pump inhibitor (PPI)"
label(df$GAS) <- "Receipt of gastric acid suppressant (GAS)"
label(df$anyGAS) <- "Receipt of PPI or GAS" 
label(df$LAX) <- "Receipt of laxative"
label(df$ABX) <- "Receipt of antibiotic"
label(df$AMR) <- "Receipt of antimicrobial"
```

# Create Table 1 showing frequencies for all predictors, stratified by hospital-acquired C diff diagnosis

```{r table1 package}
# Create Table 1 with all predictors, stratified by HA-CDIF diagnosis 
(t1 <- table1::table1(~ . | HACDIF, 
                      data = df,
                      topclass = "Rtable1-shade",
                      footnote = c("PPI: proton pump inhibitors", "GAS: antacids, histamine H2-receptor inhibitors", "anyGAS: proton pump inhibitors, antacids, or histamine HA2-receptor inhibitors", "LAX: laxatives and cathartics, local/rectal laxatives", "ABX: antibiotics", "AMR: antimicrobials including antibiotics, antifungals, antiinfectives, antiparasitics, antivirals"),
                      rowlabelhead = "Positive HA C. difficile infection"))

# Re-create Table 1 with additional column for p-test of Chi Square Test testing independence between HACDIF and predictor(s) 
# Create function to compute p-vaue for categorical (or continuous) variables
pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- chisq.test(table(y, g))$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}

# Re-run table1 and supply p-value function in the extra.col list argument 
(t1.b <- table1::table1(~ . | HACDIF, 
                      data = df, 
                      topclass = "Rtable1-shade", 
                      extra.col = list(`P-value` = pvalue),
                      footnote = c("PPI: proton pump inhibitors", "GAS: antacids, histamine H2-receptor inhibitors", "anyGAS: proton pump inhibitors, antacids, or histamine HA2-receptor inhibitors", "LAX: laxatives and cathartics, local/rectal laxatives", "ABX: antibiotics", "AMR: antimicrobials including antibiotics, antifungals, antiinfectives, antiparasitics, antivirals"),
                      rowlabelhead = ""))

# Attempts to Include Footnotes: 
#kableExtra::footnote(knitr::kable(t1.b, "html"), number = c("PPI: proton pump inhibitors", "GAS: antacids, histamine H2-receptor inhibitors", "anyGAS: proton pump inhibitors, antacids, or histamine HA2-receptor inhibitors", "LAX: laxatives and cathartics, local/rectal laxatives", "ABX: antibiotics", "AMR: antimicrobials including antibiotics, antifungals, antiinfectives, antiparasitics, antivirals", "For those not captured in ABX therapeutic class, cross-validated with a list of specific ABX and AMR names"), number_title = "Footnotes:", threeparttable = T, title_format = c("bold", "italic"))

#kableExtra::add_footnote(knitr::kable(t1.b, "html"), label = c("PPI: proton pump inhibitors", "GAS: antacids, histamine H2-receptor inhibitors", "anyGAS: proton pump inhibitors, antacids, or histamine HA2-receptor inhibitors", "LAX: laxatives and cathartics, local/rectal laxatives", "ABX: antibiotics", "AMR: antimicrobials including antibiotics, antifungals, antiinfectives, antiparasitics, antivirals", "For those not captured in ABX therapeutic class, cross-validated with a list of specific ABX and AMR names"), notation = "number", threeparttable = T)
```

# Information for Table 1 footnotes:

-   **PPI:** proton pump inhibitors

-   **GAS:** antacids, histamine H2-receptor inhibitors

-   **anyGAS:** proton pump inhibitors, antacids, or histamine HA2-receptor inhibitors

-   **LAX:** laxatives and cathartics, local/rectal laxatives

-   **ABX\*:** antibiotics

-   **AMR\*:** antimicrobials including antibiotics, antifungals, antiinfectives, antiparasitics, antivirals

*\*For those not captured in ABX therapeutic class, cross-validated with a list of specific ABX and AMR names.*

## Other Important Definitions

-   For patients with positive CDI culture, receipt of medications were coded as those received from time of admission to 1 day before CDI culture was ordered. For example, if a patient was admitted on 03-01-2021 and had a positive CDI culture ordered on 03-15-2021, his/her receipt of medications through 03-14-2021 were considered.

-   For patients without a CDI culture (AKA no CDI diagnosis), receipt of medications were coded as those received from time of admission to discharge. For example, if a patient without HA-CDI was admitted on 03-01-2021, all medications until his/her discharge time were considered.

-   Hospital-acquired CDI was defined as an abnormal/positive C difficile culture ordered greater than 48 hours after time of admission. In our analyses, if the difference in time between `CDI_order_time` and `admit_time` is greater than 48 hours, then `HACDIF` was coded as 1 or yes. This is what is commonly used in the literature and medical field.
