---
title: "vars_for_logreg_model_CDIrisk"
author: "Steph Reynolds"
date: "`r format(Sys.time(), '%b %d, %Y  %H:%M %p')`"
output:
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Background

## Project 
MINDSCAPE: Modeling of infectious network dynamics for surveillance, control and prevention enhancement

## Description 
This file reads in necessary data and identifies variables of interest for building a multivariate logistic regression model to evaluate C diff risk among hospitalized patients. 

# Load packages 
```{r Load packages}
# Load required packages 
#library(here)
library(readr)
library(tidyverse)
library(psych)
library(lubridate)
library(tableone)
library(kableExtra)
```

# Import data (only first 5000 lines)
```{r Read in data}
dm <- read_delim("/Users/sreynolds2/Downloads/ucsf_data_pull_11.08.21/demographics and event table 11.08.21.csv", delim = "|", n_max = 5000)

lab <- read_delim("/Users/sreynolds2/Downloads/ucsf_data_pull_11.08.21/lab results 11.08.21.csv", delim = "|", n_max = 5000)

dx <- read_delim("/Users/sreynolds2/Downloads/ucsf_data_pull_11.08.21/covid diagnoses 11.08.21.csv", delim = "|", n_max = 5000)

med_admin <- read_delim("/Users/sreynolds2/Downloads/ucsf_data_pull_11.08.21/med admin table 11.08.21.csv", delim = "|", n_max = 5000)

```

# Demographics and events table
```{r Demographics and events table - clean and generate categorical frequency table}
# Clean `dm` data
dm <- dm %>% 
  select(deid_enc_id, age, sex, pat_race, ethnicity, smoking, BMI, covid_pos, readmit, HAI_pathogen, HAI_type, HAI_first_date, ED_dispo, present_source, admit_time) %>% 
  rename(ID = deid_enc_id, 
         race = pat_race) %>% 
  mutate_at(c("sex", "race", "ethnicity", "smoking", "covid_pos", "readmit", "ED_dispo", "present_source", "HAI_type"), factor) %>% 
  mutate_at(c("admit_time", "HAI_first_date"), ymd_hms) %>% 
  mutate_at("BMI", as.numeric)

# Create frequency table for `dm`
CreateCatTable(data = dm, vars = c("sex", "race", "ethnicity", "smoking", "covid_pos", "readmit", "ED_dispo", "present_source", "HAI_type"))

```

# Diagnoses table 
```{r Diagnoses table - clean and generate categorical frequency table}
# Clean `dx` data
dx <- dx %>% 
  select(deid_enc_id, dx_group, DX_NAME, present_on_admit, severity_of_dx, co_morbidity) %>% 
  mutate_at(c("dx_group", "DX_NAME", "present_on_admit", "severity_of_dx", "co_morbidity"), factor)

# Create frequency table for `dx` 
CreateCatTable(data = dx, vars = c("dx_group", "DX_NAME", "present_on_admit", "severity_of_dx", "co_morbidity"))

```

# Lab results table
```{r Lab results table - clean and generate categorical frequency table}
# Clean `lab` data
lab <- lab %>%
  select(deid_enc_id, COMMON_NAME, ORD_VALUE, Abnormal, order_time, RESULT_TIME, lab_concept) %>% 
  mutate_at(c("COMMON_NAME", "Abnormal", "lab_concept"), factor) %>% 
  mutate_at("ORD_VALUE", as.numeric)

# Create frequency table for `lab` 
CreateCatTable(data = lab, vars = c("COMMON_NAME", "Abnormal", "lab_concept"))

```

# Med admin table
```{r Med admin table - clean and generate categorical frequency table}
# Clean `med_admin` table 
med_admin <- med_admin %>% 
  select(deid_enc_id, name, thera_class, pharm_class, pharm_subclass, taken_time) %>% 
  rename(ID = deid_enc_id, 
         med_name = name) %>%
  mutate_at(c("med_name", "thera_class", "pharm_class", "pharm_subclass"), factor)

# Create frequency table for `med_admin`
CreateCatTable(data = med_admin, vars = c("med_name", "thera_class", "pharm_class", "pharm_subclass")) 

```

# Save all data 
```{r Save data as .csv files}
write_csv(dm, file = "/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/clean/dm.csv")

write_csv(dx, file = "/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/clean/dx.csv")

write_csv(lab, file = "/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/clean/lab.csv")

write_csv(med_admin, file = "/Users/sreynolds2/Documents/GitHub/MS-CDI_Risk/MS-CDI_Risk/data/clean/med_admin.csv")

```


# Identify variables of interest in datasets to include in logistic regression model 

*The following are common parameters / risk factors to include in C diff risk assessment model, and where to find them in the datasets:*

* Age>= 65 --> dm$age
* Prior hospital admission in 30 days --> dm$readmit == "Yes"
* Admission from ED --> dm$ED_dispo == "Admit" OR dm$present_source == ??
* Diabetes --> lab$lab_concept == "diabetes" OR dx$dx_group == "DIABETES MELLITUS"
* Receipt of antibiotics in last 30 days --> med_admin$thera_class %in% c("ANTIBIOTICS", "ANTIFUNGALS", "ANTIINFECTIVES/MISCELLANEOUS", "ANTIPARASITICS", "ANTIVIRALS")
  + NOTE: Doesn't include outpatient abx, only those during hospitalization
* Receipt of proton pump inhibitor (PPI) --> med_admin$pharm_class == "PROTON-PUMP INHIBITORS"
* Receipt of gastric acid suppressants, laxatives, etc. --> med_admin$pharm_class %in% c("ANTACIDS", "LAXATIVES AND CATHARTICS", "LAXATIVES, LOCAL/RECTAL")

#levels(as.factor(med_admin$thera_class))

#med_admin %>% filter(thera_class == "GASTROINTESTINAL") %>% select(pharm_class) %>% unique()

# Notes / Next Steps 

* Create multivariate logistic regression model, divide data into 3 chunks/time periods (e.g. 2019, 2020, 2021)
  + 1 chunk used to train model, the other 2 used to evaluate model performance (e.g. 2019 chunk used to train model, while 2020-2021 used to test model performance)
* Evaluate "model creep" or how much the optimized model changes over time? 
* Question: What proportion of C diff cases occurred 2+ days after admission vs. within 2 days of admission? 

```{r, eval = F}
# Add column to `dm` to indicate whether CDIFF dx was within 2 days or > 2 days of admission
dm <- dm %>% 
  mutate(HAI_first_date = as.POSIXct(HAI_first_date, format = "%Y-%m-%d %H:%M:%OS3"),
         admit_time = as.POSIXct(admit_time, format = "%Y-%m-%d %H:%M:%OS3")) %>% 
  mutate(days_admit_to_dx = difftime(HAI_first_date, admit_time, units = "hours"))
# need to fix datetime format to include millisecs??? idk how to do this?
```


